### 2015-04-09 Presentation/Presentation.tex
Presentation template
Added template for the presentation

% Just notes
%\documentclass[notes=only]{beamer}

% Both slides and notes
\documentclass[notes]{beamer}

% Just slides
% \documentclass{beamer}

\usetheme{default}

% Used to do math and such (I think...)
\usepackage{amsmath}
\usepackage{amssymb}

% Used to color text (for todos)
\usepackage{xcolor}

\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{array}

% Used to embed pdfs from yEd and other sources
\usepackage{graphicx}

% Used to make code listings
\usepackage{listingsutf8}
% Used to adjust the page margin
\usepackage{geometry}
% Used to make code show up in multiple columns to fit more on a printed page.
\usepackage{multicol}

% Used to make tables, figures, etc show up where I want them.
\usepackage{float}

\usepackage{array}

%import figures from photoshop
\usepackage{epstopdf}

% Deal with backwards quotes because evidently Latex doesn't know better.
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}

% multiple rows
\usepackage{multirow}

% used for lettered bullet points
\usepackage{enumerate}

% http://tex.stackexchange.com/questions/137022/how-to-insert-page-number-in-beamer-navigation-bars
% Add page numbers
\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\begin{document}

\title{Deep Learning for Automatic Speech Recognition}
\subtitle{CSCI/LING-5832: Natural Language Processing}
\author{Ryan Hartsfield \and Garrett Lewellen}
\date{April 23, 2015}

\frame{\titlepage}

\begin{frame}
	\frametitle{Outline}

	\begin{enumerate}
		\item \textcolor{blue}{Introduction}
		\item Gaussian Mixture Models
		\item Convolutional Neural Networks
		\item Deep Neural Networks
		\item Experimental Results
		\item Conclusions
	\end{enumerate}
\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}
	\frametitle{Outline}

	\begin{enumerate}
		\item Introduction
		\item \textcolor{blue}{Gaussian Mixture Models}
		\item Convolutional Neural Networks
		\item Deep Neural Networks
		\item Experimental Results
		\item Conclusions
	\end{enumerate}
\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}
	\frametitle{Outline}

	\begin{enumerate}
		\item Introduction
		\item Gaussian Mixture Models
		\item \textcolor{blue}{Convolutional Neural Networks}
		\item Deep Neural Networks
		\item Experimental Results
		\item Conclusions
	\end{enumerate}
\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}
	\frametitle{Outline}

	\begin{enumerate}
		\item Introduction
		\item Gaussian Mixture Models
		\item Convolutional Neural Networks
		\item \textcolor{blue}{Deep Neural Networks}
		\item Experimental Results
		\item Conclusions
	\end{enumerate}
\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}
	\frametitle{Outline}

	\begin{enumerate}
		\item Introduction
		\item Gaussian Mixture Models
		\item Convolutional Neural Networks
		\item Deep Neural Networks
		\item \textcolor{blue}{Experimental Results}
		\item Conclusions
	\end{enumerate}
\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}
	\frametitle{Outline}

	\begin{enumerate}
		\item Introduction
		\item Gaussian Mixture Models
		\item Convolutional Neural Networks
		\item Deep Neural Networks
		\item Experimental Results
		\item \textcolor{blue}{Conclusions}
	\end{enumerate}
\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}

\end{frame}

\begin{frame}
	\begin{center}
		Questions?
	\end{center}
\end{frame}

\end{document}